---
title: "GetControls"
author: "Zongchao Liu"
date: "8/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(arsenal)
library(knitr)
```

# Data preparation

```{r, message = FALSE, warning=FALSE}
# existing list
subjects = read_csv('./selected_subjects.csv') %>%
  .[,-1] %>%
  rename("id" = "x")

demographic = read_csv('./demographic/ABCDdata_demographics.csv') %>%
  mutate(ID = str_remove(ID, "_")) %>% 
  filter(is.na(CBCL_Attention_T) == FALSE) %>%
  mutate(group = ifelse(ID %in% subjects$id ,"adhd", "control")) # in total num of adhd == 680 in demographics dataset

# split the whole data imto 3 parts
# att_above_70 for picking groups with socre >= 70
# att_50_55 for picking groups with socre >= 50 & <= 55
# att_56_69 for picking groups with socre >= 56 & <= 69

att_above_70 = demographic %>%
  filter(CBCL_Attention_T >= 70) %>%
  select(ID, Age, SexF, CBCL_Attention_T, CognitionTotal) %>%
  mutate(selection = ifelse(ID %in% subjects$id, "exist", "new"))

att_50_55 = demographic %>%
  filter(CBCL_Attention_T >= 50 & CBCL_Attention_T <=55) %>%
  select(ID, Age, SexF, CBCL_Attention_T, CognitionTotal)

att_56_69 = demographic %>%
  filter(CBCL_Attention_T >= 56 & CBCL_Attention_T <= 69) %>%
  select(ID, Age, SexF, CBCL_Attention_T, CognitionTotal)
```

# Define the function for matching(for one case): 

```{r}
# define a sampler
sampler = function(df = att_above_70, idx, pool = att_50_55, drop = TRUE, age_range = 12){
  
  # drop = True: initially drop rows with missing IQ measures(only 17 subjects are missing), so the 70+ group won't have any NA in IQ.
  # age_range: to control the level of restrictions on age matching
  # df: group with score > 70
  # idx: index of one subject in `df`
  # pool: the pool where we want to find the controls, can be either 50~55 or 56~69
  
  if(drop){df = df %>% filter(is.na(CognitionTotal) == FALSE)}
  
   case = df[idx, ]
  
  # start matching
  search = pool %>% 
    drop_na()
  
  if(is.na(case$CognitionTotal)){
    search = search %>%
      filter(SexF == case$SexF) %>%
      filter(abs(Age-case$Age) == min(abs(Age-case$Age)))
  }else{search = search %>%
    filter(SexF == case$SexF) %>%
    filter(abs(CognitionTotal-case$CognitionTotal) == min(abs(CognitionTotal-case$CognitionTotal))) %>%
      filter(abs(Age-case$Age) < min(abs(Age-case$Age)) + age_range)}
  
  if(nrow(search) == 0){matched = case %>% mutate(selection = ifelse(ID %in% subjects$id, "exist", "new"))}
  if(nrow(search) == 1){matched = search %>% mutate(selection = ifelse(ID %in% subjects$id, "exist", "new"))}
  if(nrow(search) > 1){matched = search[sample(1:nrow(search), size = 1), ] %>% mutate(selection = ifelse(ID %in% subjects$id, "exist", "new"))}
  
  return(matched) # return all available matched controls
}
```

# Define the function for matching(for one group): 

```{r}
## procedure 1
get_control_1 = function(df = att_above_70, drop = TRUE){
  if(drop){df = df %>% filter(is.na(CognitionTotal) == FALSE)}
  controls_1 = data.frame()
  
  for (idx in 1:nrow(df)) {
  new_row = sampler(pool = att_50_55, idx = idx, drop = TRUE)
  #print(idx)
  controls_1 = rbind.data.frame(controls_1, new_row)
  }
  return(controls_1)
}

## procedure 2
get_control_2 = function(df = att_above_70, drop = TRUE){
  
  if(drop){df = df %>% filter(is.na(CognitionTotal) == FALSE)}
  controls_2 = data.frame()
  
  for (idx in 1:nrow(df)) {
  new_row = sampler(pool = att_56_69, idx = idx, drop = TRUE)
  #print(idx)
  controls_2 = rbind.data.frame(controls_2, new_row)
  }
  return(controls_2)
}
```

# get controls for the group with score == 50~55 / 56~69

```{r}
set.seed(123)
# get controls
controls_1 = get_control_1() # 70 vs 50~ 55
controls_2 = get_control_2() # 70 vs 56 ~ 69

# merge data
list = rbind.data.frame(att_above_70, controls_1, controls_2) %>%
  mutate(group = ifelse(CBCL_Attention_T >= 70, "case", "control"),
         SexF = factor(SexF)) %>%
  distinct()# this is the final list(containing both existing or new data)

# export
new_subjects = list %>% filter(selection == "new")
sum(new_subjects$ID %in% subjects$id) == 0
write.csv(new_subjects[,1], './controls/new_list.csv',row.names = FALSE)
```


# check data

## number of new / existing subjects

```{r,echo=FALSE}
table(list$selection)

list %>%
  group_by(group, SexF) %>%
  summarise(n = n()) %>% 
  kable()
```

## correlation

```{r,echo=FALSE}
cor(list[-which(is.na(list$CognitionTotal)),-c(1,3,6,7)]) %>% kable()
```


# check distribution (gender/IQ/age)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# check gender
list %>% ggplot(aes(x = CBCL_Attention_T, fill = factor(SexF), y = ..density..))+
  geom_histogram(alpha = .5) +
  geom_density(aes(x =CBCL_Attention_T, y = ..density.., color = factor(SexF) ),alpha = .4) +
  facet_grid(factor(SexF) ~. ) +
  labs(title = "Gender")

# check IQ
list %>% ggplot(aes(x = CognitionTotal , fill = group, y = ..density..)) + geom_histogram(alpha = .5) +
   geom_density(aes(x = CognitionTotal, y = ..density.., color = group ),alpha = .4) +
  facet_grid(group ~ .) +
  labs(title = "IQ")

# check age
list %>% ggplot(aes(x = Age, fill = group)) + geom_histogram(alpha = .5) +
   geom_density(aes(x = Age, y = ..density.., color = group ),alpha = .4) +
  facet_grid(group ~ .) +
  labs(title = "Age")
```


# Table

The resulting data have 11 subjects with NAs in `CognitionTotal`. This is because I didn't dropped these subjects(11 out of 292) when I selected subjects with attention score >= 70. Among the 11 subjects, 5 are existing and 6 are new.

```{r, echo=FALSE}
list %>%
  filter(is.na(CognitionTotal)) %>%
  group_by(selection) %>%
  summarise(count = n()) %>% 
  kable()
```


```{r, echo=FALSE}
table = tableby(group ~ ., list[,-c(1,6)])
summary(table, title = "Charateristics of ADHD vs Control group") %>% kable()
```


